---
title: "practical_5"
author: "Alexander Maslev, Hanxiong Wang, Minyoung Lee"
date: "2018 /03 / 23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## EXERCISE 2
Simulations: Selection by Hypothesis Testing

```{r warning=FALSE,echo=FALSE,results=FALSE}

require(MASS)
library(selectiveInference)
library(tictoc)
set.seed(11)
```

### Simulation settings

### 1) Generate MVN
we have created the matrix $X_{n*p}$ with n = 1000 and p = 10.

```{r,echo=FALSE}

p<-10
n<-1000
sigma<-diag(p)
Mu<-rep(0,p) # location vector
X<-mvrnorm(n , Mu, sigma)

```

### 2) Creating beta
$\beta$ = [3 1.5 0 2 rep(0,6)]

```{r,echo=FALSE}

beta<-c(3,1.5,0,2,rep(0,6))

```

### 3) $\hat{y}$ = $X_{n*p}\beta + \epsilon$

```{r,echo=FALSE}

e<-rnorm(n, mean = 0, sd = 1)
Y_hat<-X%*%beta+e

```

### SELEVTIVE INFERENCE
We have done the stepwise regression and model selection with ForwardStop rule, AIC and BIC

### a)Stepwise regression with forwardStop 
```{r}

fsfit1<-fs(X,Y_hat)
out1<-fsInf(fsfit1)
forwardStop(out1$pv,alpha=0.05)

```
We did forward stepwise regression by fs(), then inference for forward stepwise regression by fsInf() and we selected model based on forward stop rule.\

We tried with 7 different values for type one error($\alpha$=[0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4])
```{r}

forwardStop(out1$pv,alpha=0.1)
forwardStop(out1$pv,alpha=0.15)
forwardStop(out1$pv,alpha=0.2)
forwardStop(out1$pv,alpha=0.25)
forwardStop(out1$pv,alpha=0.3)
forwardStop(out1$pv,alpha=0.35)
forwardStop(out1$pv,alpha=0.4)

```
Until $\alpha$=0.35, the model selection procedure stopped in 3rd step which includes variable 1, 4 and 2 in each step. But if we increase the type one error to 0.4 then we includes two more variables which are variable 10 and 9.

### b) Stepwise regression with AIC and BIC
```{r}
fsfit2<-fs(X,Y_hat)
out2<-fsInf(fsfit2,type="aic",mult = 2,alpha=0.05)
out2


fsfit3<-fs(X,Y_hat)
out3<-fsInf(fsfit2,type="aic",mult = log(n),alpha=0.05)
out3

```
We did the forward stepwise regression by fs(), then inference for forward stepwise regression by fsInf() and we selected model based on AIC and BIC.\
First result is the model selection based on AIC. By this rule, the forward stepwise regression stops at 5th step with variable 1, 4, 2, 10 and 9 included. This is same as forwardstop rule with $\alpha$ = 0.4.
Second result is the model selection based on BIC. By this rule, the forward stepwise regression stops at 3rd step same as forwardstop rule with $\alpha$ = 0.05. In this case, we includes variable 1, 4 and 2 in the model.

### c)Time takes for model selection procedure 
```{r,echo=FALSE}

tic()
fsfit1<-fs(X,Y_hat)
out<-fsInf(fsfit1)
forwardStop(out$pv,alpha=0.10)
toc()

tic()
fsfit2<-fs(X,Y_hat)
out2<-fsInf(fsfit2,type="aic",mult = 2,alpha=0.05)
toc()

tic()
fsfit3<-fs(X,Y_hat)
out3<-fsInf(fsfit2,type="aic",mult = log(n),alpha=0.05)
toc()


```
Forward Stepwise model selection in our case has 55 combination of the models to check. And It takes approximately 2sec for this process. If we wanted to do the exhaustive search, we need 2^p-1 = 1023 models to check. Then it will probably takes 40sec to do the exhaustive search. With just 10 variables, It is 40 sec, but if we have high dimension, time for the model selection process will grow rapidly.

## EXERCISE 2
Real Data Application : Zambia data set

```{r}
load("malnutrion_zambia_cleaned.Rda")
#str(data_zambia)

Y_zambia<-data_zambia[,1]
X_zambia<-as.matrix(data_zambia[,c(2:7,21,22,24)])


fitzambia<-fs(X_zambia,Y_zambia)
out_zambia<-fsInf(fitzambia)
out_zambia
forwardStop(out_zambia$pv,alpha=0.05)
out_zambia_aic<-fsInf(fitzambia,type="aic",mult = 2,alpha=0.05)
out_zambia_aic
out_zambia_bic<-fsInf(fitzambia,type="aic",mult = log(n),alpha=0.05)
out_zambia_bic
```
First output shows the result based on Forwardstop rule. The process stopped at 2nd step and variable 1 and 5 is included. Second output shows the result based on AIC rule. The process stopped at 9th step and variable 1, 5, 8, 7, 6, 3, 2, 9 and 4 is included in a sequence. Last output with BIC rule stopped at 5th step. Thus variable 1, 5, 8, 7 and 6 is included sequentially.



