---
title: "Classify Tree"
author: "Hanxiong.Wang"
date: "April 9, 2018"
output: html_document
---
## Regression Trees
We need to ge the library and load the data here is the way we lode the data.
To use the function tree grow we inport the library "repart"
```{r}

############## Regression Trees
library("rpart", lib.loc="C:/Program Files/R/R-3.4.4/library")
data <- iris
n <- nrow(data)
index <- sample(1:n, size=0.6*n)

```
## Split data
Here we split the data randomly and 1/3 fro test and 2/3 for the train data. After that we build the model b using rpart function.and the plot is below
```{r}
# Split data

data_train<- data[index,]
data_test<- data[-index,]
str(data_train)
rpart.control(minsplit = 20)
model <- rpart(formula = Species~.,data = data_train,minbucket = 1)
plot(model)
plot(model, uniform = TRUE,compress = TRUE,margin = 0.11)
text(model, splits = TRUE)

```

## tree prune
Here we need to prune the tree for better prediction and model.And here we use the prune the tree by using the "prune" function and the result is showing below.

```{r}

# tree prune
tree_prune <- prune(tree = model,cp = 0.1)
plot(tree_prune,uniform = TRUE,compress = TRUE,margin = 0.3)
text(model, splits = TRUE)


```


## Predict
Here we use the model we build to make a prediction and we can get a out of simple error by using the test data.And here is the test error rate below.
```{r}
# Predict
pred <- predict(object = model,newdata = data_test,type = "class")
real <- data_test[,5]
# How many wrong 
err_pred <- which(pred!=real)
```

## Data Zambia
Similar with the exercise before we deal with the Data Zambia. And follow the similar way to do that.

## Split data
Here we split the data randomly and 1/3 fro test and 2/3 for the train data. After that we build the model by using rpart function.
```{r}
############## Data Zambia
setwd("D:/Gen?ve Universit?/High Dimention Data/practical_6")
load("malnutrion_zambia_cleaned.Rda")
data_zambia <- data_zambia[,c(1:7,21,22,24)]
n_2 <- nrow(data_zambia)
index <- sample(1:n_2, size=0.6*n_2)


data_train_2<- data_zambia[index,]
data_test_2<- data_zambia[-index,]
str(data_train_2)

model.zb <- rpart(formula = `Height for age sd`~.,data = data_train_2,minbucket = 1)
plot(model.zb)
plot(model.zb, uniform = TRUE,compress = TRUE,margin = 0.11)
text(model.zb, splits = TRUE,cex = 0.8)

```


## tree prune
Here we need to prune the tree for better prediction and model.And here we use the prune then before we prune the tree
```{r}
# tree prune
tree_prune_2 <- prune(tree = model.zb,cp = 0.1)
plot(tree_prune_2,uniform = TRUE,compress = TRUE,margin = 0.3)
text(model.zb, splits = TRUE)


```
## Predict
Here we use the model we build to make a prediction and we can get a out of simple error by using the test data.And here is the test error rate below.
Because the y is a continue variable so we only calculate the rss to see how good the model is.

```{r}
# Predict
pred_2 <- predict(object = model.zb,newdata = data_test_2)
real <- data_test_2[,1]
d    <- as.matrix(real-pred_2)
rss <- t(d)%*%d

```





