---
title: "Stepwise Regression and Selection by Hypothesis Testing"
author: "Nensi Kopellaj & Cecilia Galvan"
date: "22 03 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(selectiveInference)
library(tictoc)
```

# 2 Simulations: Selection by Hypothesis Testing

## 1

```{r}
set.seed(11)
n <- 1000
p <- 10
mu <- rep(0, length = p)
sigma <- diag(p)
X <- mvrnorm(n, mu, sigma)
colnames(X) <- c('X1','X2','X3','X4','X5','X6','X7','X8','X9','X10')
```

## 2

```{r}
b <- c(3,1.5,0,2,rep(0,6))
```

## 3

```{r}
y_hat <- X %*% t(t(b)) + mvrnorm(n, 0, 1)
```

## (a)

```{r}
forward.stepwise.reg <- fs(X,y_hat)
forward.stepwise.reg.inf <- fsInf(forward.stepwise.reg, alpha = 0.05)
forwardStop(forward.stepwise.reg.inf$pv, alpha = 0.05)

alphas <- seq(0.01,0.5,0.01)
k.hat <- c(1:length(alphas))
for(i in k.hat){
  k.hat[i] <- forwardStop(forward.stepwise.reg.inf$pv, alpha = alphas[i])
}

plot(alphas,k.hat)
```

For large alphas (bigger than 0.37), it doesn't select the right model as it chooses 5 parameters whereas 3 is the right number. For an alpha larger than 0.45, it will select a model with 6 parameters.

## (b)

```{r}
fsInf(forward.stepwise.reg, alpha = 0.05,type = 'aic')
fsInf(forward.stepwise.reg, alpha = 0.05,type = 'aic', mult = log(12))
x_data <- data.frame(X,y_hat)
regg <-lm(y_hat ~ X1 + X4 + X2 + X10 + X9 + X3, data = x_data )
AIC(regg)
```

After we added the variables following the order of the fs() function we saw that after adding the first 5 variables, on the 6th the AIC started to increase. This is coherent with the results in the first two lines.

## (c)

Number of possible models: $$ 2^p - 1 = 1023 $$

```{r}
2^10-1
```

We are in presence of too many variables to make an exhaustive search, all calculations will take several YEARS!!!

## (d)

```{r, error=FALSE}
set.seed(11)
n <- 100
p <- 150
mu <- rep(0, length = p)
sigma <- diag(p)
X <- mvrnorm(n, mu, sigma)
b <- c(3,1.5,0,2,rep(0,146))
y_hat <- X %*% t(t(b)) + mvrnorm(n, 0, 1)
forward.stepwise.reg <- fs(X,y_hat)
alphas <- c(0.01,seq(0.01,0.5,0.01))
step.hat <- matrix(ncol= length(alphas),nrow = 25)
tic
for (i in 1:25) {
  for(j in 1 : length(alphas) ){
  forward.stepwise.reg.inf <- fsInf(forward.stepwise.reg)
  step.hat[i,j] <- forwardStop(forward.stepwise.reg.inf$pv, alpha = alphas[j])
  }
}
toc
colnames(step.hat) <- c(0.01,seq(0.01,0.5,0.01))

plot(c(0.01,seq(0.01,0.5,0.01)),step.hat[1,], xlab = 'alpha', ylab = '# of predictors')

```

For an alpha error lower than 0.24 we only, and always, select the first parameter.

Here, we chose to iterate only 25 times because of computational time that is needed.

We observe that for a given alpha, the result is the same for all the 25 iterations.

The proportions of *Correct*, *Exact*, and the *Average #* will therefore depend on the value of alpha. We will compute these proportion with respect to the number of value in our grid of alphas (i.e. 51)

##### Correct model proportion (linked to the consistency of the model selection procedure)

Proportion of times the selected model contains the correct one.

```{r, eval=FALSE}
sum(step.hat[1,] > 3)/51
```

##### Exact model selected (X1,X4,X2)

Proportion of times the correct model is selected.

```{r, eval=FALSE}
sum(step.hat[1,]==3)/51
```

##### Average number of regressors

The average number of selected regressors.

```{r}
try <- as.factor(step.hat[1,])
levels(try)

(1*sum(step.hat[1,]=='1') + 
    2*sum(step.hat[1,]=='2') +
    3*sum(step.hat[1,]=='3') + 
    4*sum(step.hat[1,]=='4') +
    5*sum(step.hat[1,]=='5'))/51
```



# 3 Real Data Application

```{r}
data_zambia <- load("malnutrion_zambia_cleaned.Rda")
head(data_zambia)
str(data_zambia)
## Taking out the factors we have.
data_zambia <- data_zambia[,c(1:7,21,22,24)] 
X <- as.matrix(data_zambia[,-1])
y_hat <- as.vector(data_zambia[,1])
zambia.fs.reg <- fs(X,y_hat)
zambia.fs.inf <- fsInf(zambia.fs.reg, alpha = 0.05)
zambia_fstop <- forwardStop(zambia.fs.inf$pv, alpha = 0.05)
```

The stopping rules all show that we should stop at the second step and that the right model is the one with makes the height dependent on the Breastfeeding duration and on the height of the mother in meters.

```{r}
# AIC and BIC done with the packages
Aic_zambia <-fsInf(zambia.fs.reg , alpha = 0.05,type = 'aic')
Bic_zambia <-fsInf(zambia.fs.reg , alpha = 0.05,type = 'aic',mult = log(11))

# Manual AIC and BIC calculations of the model
Model_zambia <- lm(`Height for age sd`~`Breastfeeding duration (months)`+`Heigth mother (meter)`+`Child weight at birth (kg)` + `Wealth index factor score`+`Weight mother (kg)` + `Age of the mother (years)`+`Age of the child (months)`+ `Interval between births` + `BMI mother` ,data = data_zambia)
AIC(Model_zambia)
BIC(Model_zambia)
```

When we add the covariates sequentially to the stepwise calculations and compute the AIC and BIC manually we see that in fact after adding the 9th covariate (BMI mother) the AIC and BIC both start to increase compared to the previous model. This is compatible with the results in the first two computations where the results of Aic_zambia and Bic_zambia show that we should stop the calculations after the 9th step.
